Lyman-alpha spectra from rotating outflows

### 2018-03-26 paper/paper.tex
minor typos

The effects of either outflows or rotation on the \lya line have been
already explored.
Here e present for the first time a study of the joint effects of
galaxy outflows and rotation, two of the three major kinematic
components expected in LAEs.

@@

CLARA follows the propagation of individual photons through a neutral
Hydrogen medium characterized by its temperature, velocity field and
global optical depth.
The code assumes an homogeneous density throughout the simulated
volume.
In the current implementation we neglect the influence of dust.
Our basic models is an spherical distribution of neutral hydrogen,
an approximation commonly used in the literature, as it explains a
wide variety of observational features \citep{Ahn03,Verhamme06,Dijkstra06}.

The central element in this paper is the velocity field that captures
outflows and rotation.
Outflows are captured by a Hubble-like radial velocity profile with
the velocity magnitude increasing linearly with the radial
coordinate; the outflow model is fully characterized by $V_{\rm out}$, the
velocity at the sphere's surface.
Rotation follows a solid body rotation profile, which is fully
characterized by $V_{\rm rot}$, the linear velocity at the sphere's surface.

@@

In Table \ref{tab:values} we list the combination of \tauh,
\vrot and \vrot values used in this paper.
The range of values have some overlap with the expectations from a dwarf
galaxy with a total neutral hydrogen mass of $10^8$-$10^9$ $M_{\odot}$.
We run a total of $24$ different models.

@@

\cite{Garavito14} presented in the Appendix an analytical model that
accounts for these two features.
The basic assumption of the analytical model is that each
differential surface element on the sphere Doppler shifts the photons
that it emits
In this paper we introduce this ansatz by post-processing the results
of the outflow simulations without rotation.
This allows us to produce new \lya spectra and compare them with the
full radiative transfer solution including both outflows and
rotation.

@@

We evaluate first the effects of increasing rotational
velocity at fixed values of the optical depth and outflow velocity.
Figure \ref{fig:varying_rotation_small} summarizes its behavior 
(for largest range see Figure \ref{fig:varying_rotation}).
All plots correspond to the same outflow velocity of $\vout=50$\kms,
rows have constant optical depth and columns have constant rotational
velocity.
Each panel shows different morphologies corresponding to varying
line-of-sight viewing angle.

% Esta figura deberia tener 2 x 2 vinetas. vout=50.
% cada panel tiene una curva para theta=0, 45, 90.
% | tau=1E5, vrot=50, | tau=1E5, vrot=100 |
% | tau=1E6, vrot=50, | tau=1E6, vrot=100 | 
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation_small}
	\end{center}
	\caption{\textbf{Spectra for fixed outflow velocity:} We fixed $\vout=50$\kms. 
			 We vary \vrot increasing from left to right, \tauh increasing from top to 
			 bottom, and $\theta$ according to the legend.
	\label{fig:varying_rotation_small}}
\end{figure*}

We now evaluate the effects of increasing outflow velocity at fixed
values for the optical depth and rotational velocity.
Figure \ref{fig:varying_outflow_small} summarizes its behavior
(for largest range see Figure \ref{fig:varying_outflow}).
All plots correspond to the same rotation velocity of
$\vrot=50$\kms.

@@

Rotation alone does have an impact on the \lya line morphology
and we get the same influence in it that previous works. 
Together with the  outflows, we found that the resulting spectra is
equivalent to a non-rotating galaxy with outflows that undergoes a
Doppler Shift in velocity, weighted by  amount of photons that if
rotating would project their escape in $\hat{k}$'s direction:

@@

Figure \ref{fig:doppler_shift} displays this Doppler shift for four different 
configurations with a fixed optical depth \tauh. In each cell of the grid 
there is a \lya spectrum for a galaxy with only outflows and two \lya spectra 
for the galaxy with outflows and rotation, one obtained from the radiative 
transfer and the other by the Doppler shifting of the non-rotating galaxy.

% Esta figura deberia tener 3 x 2 viñetas. tau=10E6.
% cada panel tiene una curva para no rotation, rotation (radiative transfer)
% y rotation (doppler)
% |vrot=50, vout=5,  | vrot=100, vout=5,  |
% |vrot=50, vout=25, | vrot=100, vout=25, |
% |vrot=50, vout=50, | vrot=100, vout=50, |
\begin{figure*}
  \begin{center}
    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_6}
  \end{center}
  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=?$. 
  	We vary \vrot increasing from left to right and \vout increasing from top to 
  	bottom. The legend represents the \lya line obtained if there is no rotation 
  	(in thin line), if there is a radiative transfer of rotation and outflows 
  	(thick and clear line), and if there is a radiative transfer of only outflows,
  	but also a Doppler shift from the rotational velocity (thick and dark line). 
	    \label{fig:doppler_shift}}
\end{figure*}

In order to compare the line created with the radiative transfer process that 
involves the rotation (RT) and the line with only outflows that is Doppler 
shifted (DS), we calculate three main characteristics for each spectrum. We 
measure the line's mean velocity, standard deviation and skewness for both 
RT's and DS's approaches. We plot the results for each value in Figures 
\ref{fig:mean_line}, \ref{fig:standard_deviation} and \ref{fig:skewness}, 
respectively for $\tauh = 10^7$. We highlight that these 3 plots increase
their fit tendency with a higher optical depth. The reasons why will be 
discussed in section \ref{sec:discussion}.

@@

\subsection{Theoretical Insights}
\color{red}
Why Doppler Shift works. Talk about radiative transfer...
\color{black}
In Fig. \ref{fig:doppler} we present the spectrum of a LAE taken from to different 
sides of the galaxy. As the LAE is rotating, one side is being redshifted while 
the other is blueshifted. We see that the combined spectrum is a weighted line, 
in solid black, that combines these past two. This shows that there is in effect a 
Doppler Shift and justifies our choice to apply it after the outflows spectrum is 
obtained.

@@

Another application of this new model to the current interpretation of \lya
spectra, is that the central ($V=0$) emission of the spectra seen in the 
\lya line, is a consequence of the viewing angle of the galaxy and can be 
controlled by it. Several authors have suggested that this central emission 
is caused by radiation that escapes the galaxy without scattering. So we 
present an alternative that solves this issue. In addition, the effect of
outflows velocities \vrot in current models can modify the asymmetry of peaks
present in the \lya line. However the intensity of the valley and the width 
of the line were not easily reproduced. In this work we provide a tool to
provide these effects with values of \vout and \vrot that do not exceed the
typical LAEs' numbers found in the literature.

@@

	\item The effect of optical depth \tauh in a LAE spectrum is to induces a 
	redshift around the zero velocity in the \lya line.

@@

	\item Rotations induces a Doppler Shift in frequency of the only-outflow 
	spectrum of the LAE.

@@

In this appendix we show more \lya spectra in the Figures 
\ref{fig:varying_rotation} and \ref{fig:varying_outflow} 
with more combinations of \vrot and \vout.

@@

The rotation effect induces a Doppler shift in frequency and displaces the only-outflow spectra. As seen in Fig. \ref{fig:rotation_doppler_outflow} rotation takes the resulting spectrum with vrot$=0$ and displaces it from its central location. Then it weights the lines and they all merge into the resulting red \lya line.\\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.3\textwidth]{./figures/rotation_doppler_outflow}
	\end{center}
	\caption{\textbf{Rotations induces a Doppler shift of the only-outflow spectra:} Each of the black lines is then weighted to obtain the red line.
		\label{fig:rotation_doppler_outflow}}
\end{figure}

@@

Finally, all the spectra produced from this new LAE model is roughly consistent with observations. The consideration of adding the new rotation parameter to the standard only-outflow model found in literature is reasonable and very powerful. vrot and its consequent viewing angle $\theta$ can modify the morphology of the \lya line in new different ways by keeping the model simplified.\\

@@

	\item Rotations induces a Doppler shift in frequency of the only-outflow spectra.

>>>>

The effects of either outflows or rotation on the \lya line have been
already explored.
Here we present for the first time a study of the joint effects of
galaxy outflows and rotation, two of the three major kinematic
components expected in LAEs.

@@

CLARA follows the propagation of individual photons through a neutral
Hydrogen medium characterized by its temperature, velocity field and
global optical depth.
The code assumes an homogeneous density throughout the simulated
volume.
In the current implementation we neglect the influence of dust.
Our basic model is an spherical distribution of neutral hydrogen,
an approximation commonly used in the literature, as it explains a
wide variety of observational features \citep{Ahn03,Verhamme06,Dijkstra06}.

The central element in this paper is the velocity field that captures
outflows and rotation.
Outflows are captured by a Hubble-like radial velocity profile with
the velocity magnitude increasing linearly with the radial
coordinate; the outflows model is fully characterized by $V_{\rm out}$, the
velocity at the sphere's surface.
Rotation follows a solid body rotation profile, which is fully
characterized by $V_{\rm rot}$, the linear velocity at the sphere's surface.

@@

In Table \ref{tab:values} we list the combination of \tauh,
\vrot and \vrot values used in this paper.
The range of values have some overlap with the expectations from a dwarf
galaxy with a total neutral hydrogen mass of $10^8$-$10^9$ $M_{\odot}$.
We run a total of $27$ different models.

@@

\cite{Garavito14} presented in the Appendix an analytical model that
accounts for these two features.
The basic assumption of the analytical model is that each
differential surface element on the sphere Doppler shifts the photons
that it emits
In this paper we introduce this ansatz by post-processing the results
of the outflows simulations without rotation.
This allows us to produce new \lya spectra and compare them with the
full radiative transfer solution including both outflows and
rotation.

@@

We evaluate the effects of increasing rotational
velocity at fixed values of the optical depth and outflows velocity.
Figure \ref{fig:varying_rotation_small} summarizes its behavior 
(for a largest range see Figure \ref{fig:varying_rotation}).
All plots correspond to the same outflows velocity of $\vout=50$\kms,
rows have constant optical depth and columns have constant rotational
velocity.
Each panel shows different morphologies corresponding to varying
line-of-sight viewing angle.

% Esta figura deberia tener 2 x 2 vinetas. vout=50.
% cada panel tiene una curva para theta=0, 45, 90.
% | tau=1E5, vrot=50, | tau=1E5, vrot=100 |
% | tau=1E6, vrot=50, | tau=1E6, vrot=100 | 
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation_small}
	\end{center}
	\caption{\textbf{Spectra for fixed outflows velocity:} We fixed $\vout=50$\kms. 
			 We vary \vrot increasing from left to right, \tauh increasing from top to 
			 bottom, and $\theta$ according to the legend.
	\label{fig:varying_rotation_small}}
\end{figure*}

We evaluate the effects of increasing outflows velocity at fixed
values for the optical depth and rotational velocity.
Figure \ref{fig:varying_outflow_small} summarizes its behavior
(for a largest range see Figure \ref{fig:varying_outflow}).
All plots correspond to the same rotation velocity of
$\vrot=50$\kms, rows have constant optical depth and columns 
have constant outflows velocity.

@@

Rotation alone does have an impact on the \lya line morphology
and we obtain the same influence as in the literature.
Together with the outflows, we found that the resulting spectra is
equivalent to a non-rotating galaxy with outflows that undergoes a
Doppler Shift in velocity, weighted by the amount of photons that if
rotating would project their escape in $\hat{k}$'s direction:

@@

Figure \ref{fig:doppler_shift} displays this Doppler shift for six different 
configurations with a fixed optical depth \tauh. In each cell of the grid 
there is a \lya spectrum for a galaxy with only outflows and two \lya spectra 
for the galaxy with outflows and rotation: one obtained from the radiative 
transfer simulation and the other obtained from the Doppler shifting of the 
non-rotating galaxy.

% Esta figura deberia tener 3 x 2 viñetas. tau=10E6.
% cada panel tiene una curva para no rotation, rotation (radiative transfer)
% y rotation (doppler)
% |vrot=50, vout=5,  | vrot=100, vout=5,  |
% |vrot=50, vout=25, | vrot=100, vout=25, |
% |vrot=50, vout=50, | vrot=100, vout=50, |
\begin{figure*}
  \begin{center}
    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_6}
  \end{center}
  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=90^\circ$. 
  	We vary \vrot increasing from left to right and \vout increasing from top to 
  	bottom. The legend represents the \lya line obtained if there is no rotation 
  	(in thin line), if there is a radiative transfer of rotation and outflows 
  	(thick and clear line), and if there is a radiative transfer of only outflows,
  	but also a Doppler shift from the rotational velocity (thick and dark line). 
	    \label{fig:doppler_shift}}
\end{figure*}

In order to compare the line created with the radiative transfer (RT) process that 
involves the rotation and the line with only outflows that is Doppler 
shifted (DS), we calculate three main characteristics for each spectrum. We 
measure the line's mean velocity, standard deviation and skewness for both 
RT's and DS's approaches. We plot the results for each value in Figures 
\ref{fig:mean_line}, \ref{fig:standard_deviation} and \ref{fig:skewness}, 
respectively for $\tauh = 10^7$. We highlight that these 3 plots increase
their fit tendency with a higher optical depth. The reasons why will be 
discussed in section \ref{sec:discussion}.

@@

\subsection{Theoretical Insights}
\color{red}
Why Doppler Shift works. Talk about radiative transfer...
\color{black}
In Fig. \ref{fig:doppler} we present the spectrum of a LAE taken from to different 
sides of the galaxy. As the LAE is rotating, one side is being redshifted while 
the other is blueshifted. We see that the combined spectrum is a weighted line, 
in solid black, that combines these past two. This shows that there is in fact a 
Doppler Shift and then it justifies our choice to apply it after the outflows spectrum is 
obtained.

@@

Another application of this new model to the current interpretation of \lya
spectra, is that the central ($V=0$) emission of the spectra seen in the 
\lya line, is a consequence of the viewing angle of the galaxy and can be 
controlled by it. Several authors have suggested that this central emission 
is caused by radiation that escapes the galaxy without scattering. So we 
present an alternative that solves this issue. In addition, the effect of
outflows velocities \vrot in current models can modify the asymmetry of peaks
present in the \lya line. However the intensity of the valley and the width 
of the line were not easily reproduced. In this work we provide a tool to
provide these effects with values of \vout and \vrot that do not exceed the
typical LAEs' values found in the literature.

@@

	\item The effect of optical depth \tauh in a LAE spectrum is to induce a 
	redshift around the zero velocity in the \lya line.

@@

	\item Rotations induces a Doppler Shift in frequency of the only-outflows 
	spectrum of the LAE.

@@

In this appendix we show more \lya spectra in the Figures 
\ref{fig:varying_rotation} and \ref{fig:varying_outflow} 
with the complete range of combinations of \vrot and \vout.

@@

The rotation effect induces a Doppler shift in frequency and displaces the only-outflows spectra. As seen in Fig. \ref{fig:rotation_doppler_outflow} rotation takes the resulting spectrum with vrot$=0$ and displaces it from its central location. Then it weights the lines and they all merge into the resulting red \lya line.\\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.3\textwidth]{./figures/rotation_doppler_outflow}
	\end{center}
	\caption{\textbf{Rotations induces a Doppler shift of the only-outflows spectra:} Each of the black lines is then weighted to obtain the red line.
		\label{fig:rotation_doppler_outflow}}
\end{figure}

@@

Finally, all the spectra produced from this new LAE model is roughly consistent with observations. The consideration of adding the new rotation parameter to the standard only-outflows model found in literature is reasonable and very powerful. vrot and its consequent viewing angle $\theta$ can modify the morphology of the \lya line in new different ways by keeping the model simplified.\\

@@

	\item Rotations induces a Doppler shift in frequency of the only-outflows spectra.
